---
output: github_document
---

```{r setup, include=FALSE, message=FALSE, warning=TRUE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.align = 'center', fig.path = "man/figures/")
```

# rd2markdown

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/rd2markdown)](https://cran.r-project.org/package=rd2markdown)
(Build status badge placeholder)
(Coverage badge palceholder)
(Total download badge)

# Overview

The `rd2markdown` package provides a way to conveniently convert .Rd package documentation files into markdown files. Documentation can be acquired in various ways, for instance using the documentation page of an already installed package, or by fetching it directly from the source file. The package aims to help other developers in presenting their work, by converting documentation of their packages to markdown files that later can be funneled into a report, README, and any others. Therefore there are countless different scenarios where this tool might be useful. We plan to drive the development of `rd2markdown` in a way it will address the needs of the R community.

The core logic behind the whole conversion is treating each rd tag as a separate class. We take the documentation object and parse it layer by layer like an onion. Thanks to that, each tag receives a dedicated treatment so it can be presented basically according to the demand. On top of that, it makes the package way more extensible, as adding the support of a new tag is essentially just implementing another method. 

# Installation

Install the development version from GitHub:
```{r eval = FALSE}
remotes::install_github("Genentech/rd2markdown")
```
or latest CRAN version

```{r eval = FALSE}
install.packages("rd2markdown")
```

# Examples

Below are some basic examples that show the core functionalities of the `rd2markdown`, its strong sides and possibilities it provides for the R community.

```{r}
rd_file_example <- system.file("examples", "rd_file_sample.Rd", package = "rd2markdown")
rd <- rd2markdown::get_rd(file = rd_file_example)
cat(rd2markdown::rd2markdown(rd))


cat(rd2markdown::rd2markdown(topic = "rnorm", package = "stats", fragments = c("description", "usage")))
```


To see the examples showing the complexity of the package, please take a look at the vignette attached in this package `The rd2markdown package intro`. You will find there not only examples for all exported functions with various combinations of parameters, but also the rationale for this package, why it was created, and what logic it follows.

# Plans for the future development

Below is the list of our plans for the nearest future`;

1. Implement a mechanism that discoverers custom functions that are not stored as tags like `\pkgfun{}`.
2. Deploy the package to CRAN
3. Fix bugs risen by the community. 