---
title: "The rd2markdown package intro"
author: "Szymon Maksymiuk"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{The rd2markdown package intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The rd2markdown package

Markdown has become a staple markup syntax and is commonly used as an interface
in modern developer tools. Markdown is used for READMEs, for static content
generation and as a rich text interchange format for language servers. As these
use cases become more common, it becomes increasingly important that R provides
an interface to its documentation to more seamlessly integrate with these tools.

The `rd2markdown` itself is a dependency-less, base R centric package that
converts `.Rd` documentation objects into markdown files. This isn't the first
package to attempt to solve this problem, but we feel this approach is more
robust, extensible and maintainable. We aim to provide a solution that, with a
help of the R community, will satisfy requirements set by various use cases and
groups.

# Functionalities

The `rd2markdown` package provides tools to conveniently extract the
documentation objects and then convert them into standard markdown files. Right
now, those functionalities are provided by three functions exported for the
end-user. 

## rd2markdown

### Overview

The `rd2markdown` is the core of the `rd2markdown` package. This interface is
modelled after the `tools::Rd2*` family of functions, filling the gap in this
family of functions to target markdown. It is here, that the actual conversion
from `.Rd` into `.md` happens. It uses the innovative idea of treating
particular Rd tags as classes and dispatching to particular methods accordingly.
Thanks to that, most of the common tags have dedicated treatment and can be
formatted in various ways. On top of that, such a design ensures trouble-free
extensibility of the package, should new tags need to be covered. 

`rd2markdown` can work in various types. It accepts .Rd objects extracted with
`get_rd` option but also can directly fetch documentation topics based on the
file path or the package. `rd2markdown` will leverage this functionality when
needed to flexibly convert provided Rd files provided as text, file paths or
help aliases. It is worth pointing out that the actual output of `rd2markdown`
function is a character vector of length one that uses special signs to format
the document. Use `writeLines` function to create an actual markdown file using
this output.

At the beginning, lets set up the paths for examples files we are going to use
throughout the vignette.

```{r}
rd_files <- c("rd_file_sample.Rd", "rd_file_sample_2.Rd")
rd_examples <- system.file("examples", rd_files, package = "rd2markdown")
```

### Examples

Basic usage with previously extracted documentation

```{r}
rd <- rd2markdown::get_rd(file = rd_examples[[1]])
md <- rd2markdown::rd2markdown(rd)
cat(md)
```

The content of the document object can be trimmed to particular sections with
`fragments` parameter. It corresponds to actual R package documentation sections

```{r}
rd <- rd2markdown::get_rd(file = rd_examples[[1]])
md <- rd2markdown::rd2markdown(rd, fragments = c("description", "usage"))
cat(md)
```

`rd2markdown` function can also work with a path to the file taking advantage of
the hidden parameter `file`

```{r}
md <- rd2markdown::rd2markdown(file = rd_examples[[1]])
cat(md)
```

or take documentation object directly from the package with a help of `topic`
and `package`

```{r}
md <- rd2markdown::rd2markdown(
  topic = "rnorm",
  package = "stats",
  fragments = c("description", "usage")
)

cat(md)
```

## get_rd

### Overview

The `get_rd` function actually extracts the .Rd file from the file or directly
from the package. It can be used on its own, but also `rd2markdown` function
uses it internally. The function itself is using the base function
`tools::parse_Rd` and `base::help` to extract the documentation while escaping
errors with graceful exceptions handling. 

### Examples

Extracting documentation directly from the .Rd file and `file` parameter

```{r}
rd2markdown::get_rd(file = rd_examples[[1]])
```

reaching out to the installed package to fetch documentation using `topic` and
`package` parameters while keeping `file` as `NULL`

```{r}
rd2markdown::get_rd(topic = "rnorm", package = "stats")
```

## documentation_to_markdown

### Overview

The `documentation_to_markdown` function merges both and `rd2markdown` and
`get_rd`, providing a vectorized way of parsing multiple documentation files at
once. Simply pass a vector of paths and the function will return a named list of
markdown documents created with `rd2markdown` function.

### Examples

It is enough to pass the vector of file paths to get a list of parsed markdown
files

```{r}
rd2markdown::documentation_to_markdown(rd_examples)
```

The `documentation_to_markdown` function also allows extracting only particular
parts of the documentation, using `fragments` parameter. If a given
documentation file does not have a specific section, it will become an empty
string.

```{r}
rd2markdown::documentation_to_markdown(rd_examples, fragments = "details")
```

# Summary

The `rd2markdown` package is an attempt to solve the issue of converting R
package documentation objects into readable markdown documents. It is still in
early development, so not every single aspect or nuance is covered.  Therefore,
we would like to welcome community contribution by reporting gaps by opening
issues on https://github.com/genentech/rd2markdown and describing the problem or
missing functionality. We value any type of input as we aim to provide a tool
that will be possible to use in various R applications.

